An Improved Metric for the Analysis of Swarms

### 2018-03-09 paper.tex
Minor Updates

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability1}
{Internal movement cohesion (no repulsion).\label{methods:Stability5}}

@@

\section{Internal movement and the null vector}\label{metric:StabilityNullVector}
When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions. Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed. If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect. This phenomenon is used in the identification of the emergent behaviour of \emph{area flooding}.

@@

\section{Residual internal movement (Jitter)}\label{metric:Jitter}
Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:BotStabilityT}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

@@

The distance metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:

\begin{equation}
\label{eq:SwarmPotentialMagnitude}
\psi_d(S) = \mu_d(S)\pm \sigma_d(S)
\end{equation}

@@

\section{Conclusion - metric comparison\label{metric:MagnitudeDistanceComparison}}
The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \emph{cohesion/repulsion} metric also allows the state of the swarm to be identified.

The distance metric is based upon the physical \emph{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \emph{interaction} of the agents.

The distance based metric provides an analysis of the distribution of the agents at a point in time and allows the agitation of the swarm to be assessed without considering the possible distribution of agents that the field effects \emph{could} produce.

The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential movement. This is independent of the physical distribution of the agents within the swarm. The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~\ref{additional:fieldsWork} where the physical distribution of agents may vary due to the varied field effect ranges.

In a homogeneous field effect swarm the two metrics provide a deeper evaluation of a swarm's structure and `potential'. Consider the following: the repulsion field is increased but the internal distances do not change as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution and the swarm is filling a confined area. This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}

@@

A possible adaptation to both of these areas of research would be to use the swarm perimeter detection to identify a subset of agents then reducing the field effect of those agents to increase the agent density at the perimeter. Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge. The issue with introducing heterogeneous field effects is that existing distance based metric would not allow the swarm's structure to be analysed correctly. The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify swarm stability. This is due to the metric being based upon inter-agent effects rather than physical distribution.

>>>>

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=5cm]{figures/Stability1}
{Internal movement cohesion (no repulsion).\label{methods:Stability1}}

@@

\section{Internal movement and the null vector}\label{section:StabilityNullVector}
When the two vectors (cohesion and repulsion) have magnitudes that are equal and opposite they produce a null vector, indicating that two agents are optimally spaced for a given set of conditions. Although the agents are at an optimum position when resultants are zero it does not mean the swarm is optimally distributed. If a swarm is in a confined space it is possible for an optimum position to be created where the vector magnitude is affected by a compression effect. This phenomenon is used in the identification of the emergent behaviour of \emph{area flooding}.

@@

\section{Residual internal movement (Jitter)}\label{metric:Jitter}
Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Figures~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \emph{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$ (Equation~\ref{eq:CohesionRepulsion}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

@@

The distance-based metric for the internal distribution of the agents is the pair consisting of $\mu_d(S)$, $\sigma_d(S)$. This can be written informally as:

\begin{equation}
\label{eq:SwarmDistanceMetric}
\psi_d(S) = \mu_d(S)\pm \sigma_d(S)
\end{equation}

@@

\section{Conclusion\label{metric:MagnitudeDistanceComparison}}
%% The two metrics (distance and \emph{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \emph{cohesion/repulsion} metric also allows the state of the swarm to be identified.
%% 
%% The distance metric is based upon the physical \emph{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \emph{interaction} of the agents.
The distance-based metric measures swarm structure in terms of position of agents.

The distance-based metric provides an analysis of the distribution of the agents at a point in time and allows the agitation of the swarm to be assessed. It does not take into account the possible distribution of agents that the field effects \emph{could} produce.

The \emph{cohesion/repulsion} metric provides an indication of a swarm's potential movement. This is independent of the physical distribution of the agents within the swarm. The lack of dependence on the physical distribution allows the metric to be used in heterogeneous field effect swarms~(Figure~\ref{additional:fieldsWork}) where the physical distribution of agents may vary due to the varied field effect ranges.

In a homogeneous field effect swarm the two metrics together provide a deeper evaluation of a swarm's structure and `potential'. Consider the following: the repulsion field is increased but the internal distances do not change; as a result the \emph{cohesion/repulsion metric} rises: This indicates `something' is confining the swarm's distribution. This analysis could be used in identifying effective swarm distribution for the coverage of a sensor array as discussed by Ramaithitima et al.~\cite{RWBK:15}

@@

A possible adaptation to this area of research is to use the swarm perimeter detection to identify a subset of agents then reduce the field effects (repulsion and cohesion) of those agents to increase the agent density at the perimeter. Increasing the density would allow increased coverage of any sensor arrays that the swarm was using at the swarm edge. The issue with introducing heterogeneous field effects is that the existing distance-based metric would not allow the swarm's structure to be analysed correctly. The cohesion/repulsion metric discussed in this paper would still work within this environment as a tool to identify a swarm's structure. This is due to the metric being based upon inter-agent effects rather than physical distribution.
